---
title: "simExTargId (simultaneous experiment - MS/MS target identification)"
author: "WMB Edmands"
date: '`r format(Sys.Date(), "%B %e, %Y")`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simExTargId}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The following illustrates the simExTargId works with example (.RAW) data files of 
human plasma obtained on a Thermo FT-ICR-MS:

It shows how to set up the early warning email notification process first to
monitor a dummy raw data directory on your computer's hard drive. Then
half the example .RAW data files can be moved to the dummy raw data directory
to initiate the **simExTargId** autonomous metabolomic data-analysis process.
Then another set of example .RAW data files can then be moved in to the raw data
directory to simulate a real-time collection of LC-MS data on a mass spectrometer
workstation.

```{r, include=F}
library(simExTargId)
```

# 1. Set the real-time email notifier

The email notifier must be run as a seperate monitoring function in another R session.
This is so that the raw data directory on the mass spectrometer workstation can 
be closely monitored and a stoppage in the run or any files which are much smaller
in size (greater than 3 * median absolute deviation) than the majority of the files
can be immediately identified and a warning email sent. If this function was run
from within the main **simExTargId** function there might be a delay in informing the
user of an unexpected instrument stoppage or spray quality issues for example when file-conversion,
peak-picking or another lengthy process is taking place.

```{r, eval=FALSE}
# select directory where example directories for raw files can be created and also results
# output saved
# For example the C: directory
# create a dummy raw MS1 profiling data directory 
# for example
studyName <- paste0(gsub(" ", "", format(Sys.time(), "%Y %m %d")), "_simExTargId_Example")
dummyRawDir <- paste0("C:/", studyName)
# create directory
dir.create(dummyRawDir)
# replace email address with a vector of one or more email addresses of
# people in your laboratory group
emailNotifier(rawDir=dummyRawDir, emailAddress='johndoe@emailprovider.com', 
              emailTime=5)
```

# 2. Set the main simExTargId function running

Set the main **simExTargId** function running in a different R session to the 
**emailNotifier** function. Make sure that you have the **MSConvert** software
in your path if using Windows and that you are able to successfully run the 
command `> MSConvert` in your system shell. This is necessary for the automatic 
mzXML/mzML file conversion process and **simExTargId** will not work successfully
without it.

```{r, eval=FALSE}
# simExTargId extdata directory
extdataDir <- system.file("extdata", package="simExTargId")
# list blank files
blanksRaw <- list.files(extdataDir, pattern="blank", full.names=TRUE)
# list plasma IPA extract files
samplesRaw <- list.files(extdataDir, pattern="sample", full.names=TRUE)
# covariates file
coVariates <- paste(extdataDir, "coVariates.csv", sep="/")
# illustrative metabolite database table
metabDb <- paste(extdataDir, 'exampleMetabDatabase.csv', sep='/')
# identify number of virtual cores for parallel processing using parallel package
nCores <- parallel::detectCores()

# as no qc files then use sample files twice to illustrate the
# peakMonitor function

# move the blank files into the temporary directory to start the process
blankRawCopies <- paste(dummyRawDir, basename(blanksRaw), sep="/")
file.copy(from=blanksRaw, to=blankRawCopies)
# set the file time to simulate the files having been acquired at least 5 mins
# since last modification
setTheTime <- function(fileCopy, time){
  Sys.setFileTime(fileCopy, Sys.time() - time)
}
# apply to newly copied files
sapply(blankRawCopies, setTheTime, 240)

# move the plasma samples twice first time rename as QC
# and set the file time less than 5 mins
samplesRawCopies <- paste(dummyRawDir, basename(samplesRaw), sep="/")
file.copy(from=samplesRaw, to=samplesRawCopies)
qcFiles <- gsub('sample', 'qc', samplesRawCopies)
file.rename(from=samplesRawCopies, to=qcFiles)
file.copy(from=samplesRaw, to=samplesRawCopies)

# apply to newly copied files
sapply(c(samplesRawCopies, qcFiles), setTheTime, 300)

# Start simExTargId function
simExTargId(rawDir=dummyRawDir, studyName = studyName, analysisDir='C:/',
            coVar=coVariates, metab=metabDb, nCores=nCores, ionMode='nega', 
            minFiles=3)
```

simExTargId will wait until at least five minutes after the raw data file was
last modified, to ensure that the file acquisition has completed.

After at least 3 samples of each class found in the second column of the
co-variates table, retention time alignment, grouping, zero-filling, then
pre-processing, PCA analysis, stats analysis and data-deconvolution will occur.

# 3. peakMonitor (shiny application)
A database table of previously identified metabolites can be monitored in real-time
and analytical CV% and signal attenuation affects monitored using the shiny
application **peakMonitor**
```{r, eval=FALSE}
peakMonitor(analysisDir=paste0(dummyRawDir, "_analysis/NEG/output/peakMonitor"))
```

# 4. targetId (shiny application)
During a run the output of the statistical analyses can be viewed using the 
shiny application **targetId** a zip file containing containing a .csv file for 
each statistical test after setting the thresholds can be downloaded and used 
to guide and plan further MS/MS experiments.

```{r eval=FALSE}
# this command will open the application in your web-browser
targetId(analysisDir=paste0(dummyRawDir, "_analysis/NEG/output/04.stats"))
```
